Composition-climate paper

### 2016-11-24 Response_to_reviewers.Rmd

> Lines 337-351: The authors discuss agriculture, but not timber harvesting and changes in fire regime in terms of the land-use drivers. It seems that the latter two are potentially important for some of the regional patterns. Additionally, I wonder if some of this could be discussed by looking at gains and losses relative to NLCD land-use layers. Perhaps that was in the supplement and I missed it.

We have added the sentence: *While fire and logging have played a role in changing distribution patterns in the north and northeast, they have not resulted in the exclusion or extirpation of taxa from the northern climate space, with the exception of* Tsuga.

> Figure 5: The x – axis needs more tick marks and numbers. At first, my mind assumed a linear scale, which would imply negative distances, but then I thought maybe this was some sort of log scale, which would make more sense with a distance metric. Some additional clarity is needed.

We have changed the x axis to a log10 scale and added more tick marks to the x-axis to make it clear that it is non-linear.

@@

> Table 1: Some of the taxa are listed with their scientific names and others by common names. Replace common names with scientific names

I have corrected the names so that only scientific names are displayed.

@@

> The authors should use present tense not past tense in the methods and results sections.

> The results section was not well organized. Subtitles should be used to describe the overall distribution changes, shifts along climate change, the relative importance, and interactions. The four paragraphs of interaction subsection in the results section were not concise and like bullets.

> Secondly, the materials and method section was vague and not logically articulated.
1) July temperature (Tmax) was the minimum, maximum, or mean temperature in July?
January temperature (Tmin) was the minimum, maximum, or mean temperature in January?
Annual temperature range (Tdiff) was the difference between the minimum and maximum temperature over the year? Annual precipitation (Pann) was the minimum, maximum, or mean precipitation across over the year?

We have clarified these variables: "*Estimates for mean daily July temperature ($T_{max}$), mean daily January temperature ($T_{min}$), annual temperature range ($T_{diff}$: $T_{max}$ - $T_{min}$), and annually summed daily precipitation ($P_{ann}$) . . .*"

@@

> 2)  More detailed information in the materials and method section was needed to show how the authors analyzed the overall distribution changes, climate change effects, and the relative importance of land use and climate change. It should not be only in the figure titles. For example:

> The authors used gain, loss, and continuous persistence to demonstrate the distribution changes over the last two centuries in Figure 3. However, this approach was not described in the materials and method section but was only in the title of Figure 3.

> The authors used the box-and-whisker plots to demonstrate the distribution changes along Tmax, Tmin, and Pann. Likewise, this approach was not described in the materials and method section but was only in the title of Figure 4.

> The approach used in Figure 5 was not described in the materials and method section but was only in the title of Figure 5. More detailed information should be given to better analyze the relative importance of land use change and climate change effects.

> More detailed information about the univariate density functions (PDFs) and Hellinger distances was needed to better describe the approach.

### 2016-11-24 Composition_Ecology.Rmd
Fixed the `hell_plot`, with associated comment in the response to reviewers.  Axis is now $log_{10}$, with some extra ticks.
Also added clarification as to which climate variables were used, and what they actually measure.  With response to Reviewer 3 (line 168).

Fixed the `loss_table` table, changed rownames to scientific names.

Historical and modern climate data are derived from monthly PRISM LT data for 1895 to 2014 CE [@prism2004]. The PRISM climate variables are available at 800m resolution and have been interpolated from station locations using elevation, aspect and other data.  Estimates for July temperature ($T_{max}$), January temperature ($T_{min}$), annual temperature range ($T_{diff}$), and annual precipitation ($P_{ann}$) were resampled to the 8 x 8km grid used for PLS and FIA data. These climate variables were chosen because of their relative importance as demonstrated in prior research [@iverson1998predicting] and their use in past climate reconstruction [@maiorano2013building].  Recent statistical advances have opened the possibility of obtaining temperature and precipitation estimates from historical observations as early as the 1830s [@tiptoninreview], however the behaviour of uncertainties for derived variables such as potential evapo-transpiration, which ranked highly in previous work [@iverson1998predicting] are not currently known.

@@

pander::pandoc.table(loss_table, justify = "left", style = "rmarkdown")
```

>>>>

Historical and modern climate data are derived from monthly PRISM LT data for 1895 to 2014 CE [@prism2004]. The PRISM climate variables are available at 800m resolution and have been interpolated from station locations using elevation, aspect and other data.  Estimates for mean daily July temperature ($T_{max}$), mean daily January temperature ($T_{min}$), annual temperature range ($T_{diff}$: $T_{max}$ - $T_{min}$), and annually summed daily precipitation ($P_{ann}$) were resampled to the 8 x 8km grid used for PLS and FIA data. These climate variables were chosen because of their relative importance as demonstrated in prior research [@iverson1998predicting] and their use in past climate reconstruction [@maiorano2013building].  Recent statistical advances have opened the possibility of obtaining temperature and precipitation estimates from historical observations as early as the 1830s [@tiptoninreview], however the behaviour of uncertainties for derived variables such as potential evapo-transpiration, which ranked highly in previous work [@iverson1998predicting] are not currently known.

@@

sci_names <- c("Larix", "Pinus", "Picea", "Abies", "Tsuga",
               "Thuja/Juniperus", "Populus", "Acer", "Betula", "Fagus", 
               "Ostrya/Carpinus", "Tilia", "Fraxinus", "Ulmus", "Quercus")

rownames(loss_table) <- paste0("*", sci_names, "*")

pander::pandoc.table(loss_table, justify = "left", style = "rmarkdown",
                     emphasize.rownames = FALSE)
```
